VCF file filtering by common garden site

### 2020-01-23 analysis/common-garden-allele-balancing/1_VCF_file_filtering_by_common_garden_site.Rmd
Common garden allele balancing with 33M SNP file

```{r}
colSums(!is.na(phenotypes_wide_2019))

for(i in 1:10){
site <- sites$Site[i]

@@

Plants alive EOS 2018, then EOS 2019:
BRKG: 412 down to 254
CLMB: 591 down to 496
FRMI: 273 down to 185
KBSM: 544 down to 516
KING: 305 down to 286
LINC: 437 down to 260
OVTN: 346 down to 331
PKLE: 623 down to 614
STIL: 469 down to 451
TMPL: 388 down to 373

@@

vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018j &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
vcftools --gzvcf All_chr_combined_636g_80pct_maf0.005.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &

@@

>>>>

```{r}
colSums(!is.na(phenotypes_wide_2019))
i=1
for(i in 1:10){
site <- sites$Site[i]

@@

Plants alive EOS 2018, then EOS 2019:
BRKG: 412 down to 254 (377 in 2018; 234 in 2019)
CLMB: 591 down to 496 (547 in 2018; 462 in 2019)
FRMI: 273 down to 185 (245 in 2018; 163 in 2019)
KBSM: 544 down to 516 (501 in 2018; 480 in 2019)
KING: 305 down to 286 (275 in 2018; 267 in 2019)
LINC: 437 down to 260 (400 in 2018; 247 in 2019)
OVTN: 346 down to 331 (312 in 2018; 304 in 2019)
PKLE: 623 down to 614 (579 in 2018; 577 in 2019 out of 732 individuals in 2020 genome file)
STIL: 469 down to 451 (433 in 2018; 419 in 2019)
TMPL: 388 down to 373 (351 in 2018; 341 in 2019)

@@

vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2019 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2019 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/PKLE_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/PKLE_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/BRKG_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/BRKG_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/CLMB_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/CLMB_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/FRMI_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/FRMI_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KBSM_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KBSM_2018j &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/KING_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/KING_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/LINC_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/LINC_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/OVTN_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/OVTN_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/STIL_2018 --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/STIL_2018 &
vcftools --gzvcf /home/SharedUser/PV/NatPop/33dot9MSite_Unimputed/Pvirgatum_V5_732g_33M.vcf.gz --keep /home/alice/Github/pvdiv-phenology-gxe/analysis/common-garden-allele-balancing/TMPL_2018 --freq --out /home/alice/Github/pvdiv-phenology-gxe/large-files/Site-by-year-allele-frequencies/TMPL_2018 &

@@

```{r}
snpfile <- read_csv(file = "33M_SNP_file_PLANT_ID")
pkle2019 <- read_csv(file = "PKLE_2019", col_names = "PLANT_ID")

snpfile %>%
  left_join(phenotypes_wide_2019)

metadata %>%
  filter(LIB_CLIMATE == "Y") %>%
  anti_join(snpfile)

snpfile %>%
  anti_join(metadata) # NL94SL93 is the only one not in the metadata. So what happened?

metadata2 %>%
  anti_join(snpfile) # NL94SL93 is the only one not in the 732 metadata. Ehhhh? 732 metadata has NL94 & SL93. 
# So the error isn't because I gave Taslima the wrong metadata for the 732 file.

snpfile %>%
  anti_join(pkle2019)
pkle2019 %>%
  anti_join(snpfile)
site_2019 %>%
  anti_join(snpfile) %>%
  left_join(metadata)

metadata %>%
  filter(LIB_CLIMATE == "Y")
metadata %>%
  filter(LIB_PHENO == "Y")
```
NL94SL93 is the only one not in the 732 metadata. Ehhhh? 732 metadata has NL94 & SL93.

So the error isn't because I gave Taslima the wrong metadata for the 732 file. But there are still 44 missing individuals that were planted in the field but aren't in the 732 genotype file.

I think maybe this file has the

```{r}

phenotypes_wide_2019 %>%
  anti_join(snpfile) %>%
  dplyr::select(PLANT_ID) %>%
  left_join(metadata) %>%
  arrange(LIB_PHENO) %>%
  filter(!(PLANT_ID %in% c("J661.C", "J491.A", "J385.C", "J471.A", "J243.A")))

phenotypes_wide_2019 %>%
  filter(!(PLANT_ID %in% c("J661.C", "J491.A", "J385.C", "J471.A", "J243.A", "J428.A", "J399.B", "J344.B", "J003.C", "J624.B", "J682.C", "NL94×SL93_370", "NL94×SL93_305"))) # None of these have any 2019 data.

phe_2019 <- phenotypes_wide_2019[which(rowSums(!is.na(phenotypes_wide_2019))>1),] 
# Only 630 individuals have phenotyping data in 2019. But 38 aren't in 732 file. That's 10% of data.
phe_2018 <- phenotypes_2018[which(rowSums(!is.na(phenotypes_2018))>1),] # 636 have 2018 data.
phe_csv <- phe_2018 %>%
  dplyr::select(PLANT_ID) %>%
  #anti_join(snpfile) %>%
  left_join(metadata) %>% 
  filter(LIB_PHENO == "Y") %>%
  dplyr::select(PLANT_ID:LIB_PHENO) #%>% anti_join(snpfile)

write_csv(phe_csv, path = "Individuals_needed_for_33M_SNPs.csv")
```

Individuals that are in the phenotyping panel and should be resequenced:

J428.A	IGRB	N	N	N	 Highest heterozygoisty and low-ish coverage. Fishy, but also, probaably OK. Exclude
J399.B	IHZG	N	N	N	low coverage, high het. Exclude
J344.B	IHZN	N	N	N	low coverage, high het. Exclude
J003.C	IIAH	N	N	N	J003 is a mess. Exclude
J243.A	IRUH	N	N	N	This has really bad coverage and should be excluded. Exclude
J661.C	IRXF	N	N	N		Probable Error, Plant does not match others. Exclude
J682.C	ITCI	N	N N Very high heterozygosity, coverage on the low side. Exclude
